Topics Causal Inference

### 2024-08-18 documents/paper.tex
Notes for conclusion.
Add BLP CI plot.

@@

I consider three additional DGPs: The first two DGPs increase the dimension of the CATE function, by increasing the number of arguments the function depends on.
In particular, I use
\begin{equation*}
    \tau_2(x) = \prod_{i=1}^{4}\xi(x_i),
\end{equation*}
as well as
\begin{equation*}
    \tau_3(x) = \prod_{i=1}^{8}\xi(x_i).
\end{equation*}
I keep the rest of the DGP the same. Finally I also consider a \textit{placebo} setting with
\begin{equation*}
    \tau_4(x) = 0.
\end{equation*}
The first two extensions are used to investigate the performance of the causal forest and generic machine learning approach when heterogeneity is high-dimensional.
The placebo extension is used to investigate the danger of over-fitting where machine learning algorithms might be predict a heterogeneous treatment effect when there is actually none.
A similar question is investigated in Figure 1 of~\cite{chernozhukov2023genml}.

@@

\paragraph*{Coverage, GATES and CLANS}
Table~\ref*{table:generic_ml_coverage} verifies coverage properties of the generic ML procedure for both BLP parameters $\beta_1, \beta_2$ and GATES $\mu_1, \ldots, \mu_5$.
Groups for GATES are defined by quintiles of predicted treatment effects from low ($\mu_1$) to high ($\mu_5$).
I mostly find coverage levels close to the nominal level of 95\%, although with only 40 simulation runs the accuracy is of course severely limited.
% TODO describe coverage properties, currently some seem off?
Figure~\ref*{fig:gates} illustrates the performance of the GATES procedure for the four different DGPs\@.
It shows the distribution of the estimates from the 40 simulation runs against an estimate of the true GATES from a large ($N=1,000,000$) sample draw.
While quantitatively not precise, we can still see that generic ML procedure correctly estimates almost all GATES for all four DGPs.
This is true even for the DGPs with increased heterogeneity dimension, as well as the placebo, where estimates are tightly centered around 0.

@@

@@

% Gates Figure
\begin{figure}
    \caption{Simulation Results: Generic ML --- CLANS}\label{fig:clans}

@@

>>>>

% TODO Describe causal forest?

@@

I consider three additional DGPs: The first two DGPs increase the dimension of the CATE function, by increasing the number of arguments the function depends on.
In particular, I use
\begin{equation*}
    \tau_2(x) = \prod_{i=1}^{4}\xi(x_i),
\end{equation*}
as well as
\begin{equation*}
    \tau_3(x) = \prod_{i=1}^{8}\xi(x_i).
\end{equation*}
I keep the rest of the DGP the same. Finally I also consider a \textit{placebo} setting with
\begin{equation*}
    \tau_4(x) = 0.
\end{equation*}
The first two extensions are used to investigate the performance of the causal forest and generic machine learning approach when heterogeneity is high-dimensional.
The placebo extension is used to investigate the danger of over-fitting where machine learning algorithms might be predict a heterogeneous treatment effect when there is actually none.
A similar question is analyzed in Figure 1 of~\cite{chernozhukov2023genml}.

@@

\paragraph*{Coverage, GATES and CLANS}
Table~\ref*{table:generic_ml_coverage} verifies coverage properties of the generic ML procedure for both BLP parameters $\beta_1, \beta_2$ and GATES $\mu_1, \ldots, \mu_5$.
Groups for GATES are defined by quintiles of predicted treatment effects from low ($\mu_1$) to high ($\mu_5$).
I mostly find coverage levels close to the nominal level of 95\%, although with only 40 simulation runs the accuracy is of course severely limited.
In particular, GATES and $\beta_1$ (corresponding to the ATE) have (near) correct coverage in most cases except for the very high-dimensional setting where coverage declines significantly for some parameters. Notably, $\beta_2$ has coverage of $0\%$ or close to it for almost all settings.
To investigate this further, Figure~\ref*{fig:blp} plots the distribution of the lower and upper confidence interval points for both all four DGPs and both BLP parameters.
While endpoints are to the left and right of $\beta_1$ in most cases except the high-dimensional ones, $\beta_2$ is always over-estimated.
At this point I don't have a good explanation for this. Note the true parameter is always almost equal to $1$, since we have a highly heterogeneous treatment effects (except for the placebo setting).
Hence, this could be a parameter at the boundary issue.

Figure~\ref*{fig:gates} illustrates the performance of the GATES procedure for the four different DGPs\@.
It shows the distribution of the estimates from the 40 simulation runs against an estimate of the true GATES from a large ($N=1,000,000$) sample draw.
While quantitatively not precise, we can still see that generic ML procedure correctly estimates almost all GATES for all four DGPs.
This is true even for the DGPs with increased heterogeneity dimension, as well as the placebo, where estimates are tightly centered around 0.

@@

% BLP Figure
\begin{figure}
    \caption{Simulation Results: Generic ML --- BLP Confidence Intervals}\label{fig:blp}

    \centering
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp3.png}
         \caption{Original WA DGP 3 --- 2 Covariate CATE}\label{fig_blp:dgp3}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp4.png}
         \caption{High-dimensional --- 4 Covariate CATE}\label{fig_blp:dgp4}
     \end{subfigure}

     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp5.png}
         \caption{High-dimensional --- 8 Covariate CATE}\label{fig_blp:dgp5}
     \end{subfigure}
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_blp_distribution_dgp6.png}
         \caption{Placebo}\label{fig_blp:dgp6}
     \end{subfigure}

     \floatfoot{\textbf{Notes}: This figure reports simulation results for confidence intervals of BLP parameters of the generic machine learning approach.
     For each of the four DGPs, it plots the distribution of both lower and upper confidence intervals for $\beta_1, \beta_2$ over the simulation runs.
     Lower ends are displayed red, upper blue.
     The dashed line indicates the true GATES using a large population draw. Sample size $N=10,000$ and $40$ simulation runs.
     }

@@

% CLANS Figure
\begin{figure}
    \caption{Simulation Results: Generic ML --- CLANS}\label{fig:clans}

@@

\begin{itemize}
    \item Main difficulty: Informative DGP
    \item Think of criterion relevant for decision
    \item Computation power/time to run simulations
    \item Combine WGAN and analytical CATE structure; still need to justify this.
\end{itemize}
